% \iffalse meta-comment
%
% Copyright 1994 the LaTeX3 project and the individual authors.
% All rights reserved. For further copyright information see the file
% legal.txt, and any other copyright indicated in this file.
% 
% This file is part of the LaTeX2e system.
% ----------------------------------------
% 
%  This system is distributed in the hope that it will be useful,
%  but WITHOUT ANY WARRANTY; without even the implied warranty of
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
% 
% 
% IMPORTANT NOTICE:
% 
% For error reports in case of UNCHANGED versions see bugs.txt.
% 
% Please do not request updates from us directly.  Distribution is
% done through Mail-Servers and TeX organizations.
% 
% You are not allowed to change this file.
% 
% You are allowed to distribute this file under the condition that
% it is distributed together with all files mentioned in manifest.txt.
% 
% If you receive only some of these files from someone, complain!
% 
% You are NOT ALLOWED to distribute this file alone.  You are NOT
% ALLOWED to take money for the distribution or use of either this
% file or a changed version, except for a nominal charge for copying
% etc.
% \fi
% Filename: fntguide.tex
 
\NeedsTeXFormat{LaTeX2e}[1994/12/01]
 
\documentclass{ltxguide}[1994/11/20]
 
\title{\LaTeXe{} font selection}
 
\author{\copyright~Copyright 1994, \LaTeX3 Project Team.\\
   All rights reserved.}
 
\date{1 December 1994}
 
\begin{document}
 
\maketitle
 
\tableofcontents
 
\section{Introduction}
 
This document describes the new font selection features of the \LaTeX{}
Document Preparation System.  It is intended for package writers who
want to write font-loading packages similar to |times| or |latexsym|.
This document is only a brief introduction to the
new facilities and is intended for package writers who are familiar with
\TeX{} fonts and \LaTeX{} packages.  It is \emph{neither} a user-guide
\emph{nor} a reference manual for fonts in \LaTeXe.
 
\subsection{\LaTeXe~fonts}
 
The most important difference between \LaTeX~2.09 and \LaTeXe{} is the
way that fonts are selected.  In \LaTeX~2.09, the Computer Modern fonts
were built into the \LaTeX~format, and so customizing \LaTeX{} to use
other fonts was a major effort.
 
In \LaTeXe, very few fonts are built into the format, and there are
commands to load new text and math fonts.  Packages such as |times| or
|latexsym| allow authors to access these fonts.
This document describes how to write similar font-loading packages.
 
The \LaTeXe{} font selection system was first released as the `New Font
Selection Scheme' (NFSS) in 1989, and then in release~2 in 1993.
\LaTeXe{} includes NFSS release~2 as standard.
 
\subsection{Overview}
 
This document contains an overview of the new font commands
of \LaTeX.
 
\begin{description}
 
\item[Section~\ref{Sec:text}] describes the commands for selecting fonts
   in classes and packages.  It lists the five \LaTeX{} font
   attributes, and lists the commands for selecting fonts.  It also
   describes how to customize the author commands such as |\textrm| and
   |\textit| to suit your document design.
 
\item[Section~\ref{Sec:math}] explains the commands for controlling
   \LaTeX{} math fonts.  It describes how to specify new math fonts and
   new math symbols.
 
\item[Section~\ref{Sec:install}] explains how to install new fonts into
   \LaTeX.  It shows how \LaTeX{} font attributes are turned into \TeX{}
   font names, and how to specify your own fonts using font definition
   files.
 
\item[Section~\ref{Sec:encode}] discusses text font encodings.
   It describes how to declare a new encoding and how to define
   commands, such as |\AE| or |\"|, which have different definitions in
   different encodings, depending on whether ligatures, etc.\ are
   available in the encoding.
 
\item[Section~\ref{Sec:misc}] covers font miscellania.  It describes how
   \LaTeX{} performs font substitution, how to customize fonts that are
   preloaded in the \LaTeX{} format, and the naming conventions used in
   \LaTeX{} font selection.
 
\end{description}
 
\subsection{Further information}
 
For a general introduction to \LaTeX, including the new features of
\LaTeXe, you should read \emph{\LaTeXbook},
Leslie Lamport, Addison Wesley, 2nd~ed, 1994.
 
A more detailed description of the \LaTeX{} font selection scheme is to
be found in \emph{\LaTeXcomp}, Goossens, Mittelbach and Samarin, Addison
Wesley, 1994.
 
The \LaTeX{} font selection scheme is based on \TeX, which is described
by its developer in \emph{The \TeX book}, Donald E.~Knuth, Addison
Wesley, 1986, revised in 1991 to include the features of \TeX~3.
 
Sebastian Rahtz's |psnfss| software contains the software for using a
large number of Type~1 fonts (including the Adobe Laser Writer 35 and
the Monotype CD-ROM fonts) in \LaTeX.  It should be available from the
same source as your copy of \LaTeX.
 
 
The |psnfss| software uses fonts generated by Alan Jeffrey's
|fontinst| software.  This can convert fonts from Adobe Font Metric
format into a format readable by \LaTeX, including the generation of
the |.fd| files described in Section~\ref{Sec:install}.  The
|fontinst| software should be available from the same source as your
copy of \LaTeX.
 
Whenever practical, \LaTeX{} uses the font naming scheme
described in \emph{Filenames for fonts}, Karl Berry,
\emph{TUGboat} 11(4), 1990.
 
The class-writer's guide \emph{\clsguide} describes
the new \LaTeX{} features for writers of document classes and packages
and is kept in |clsguide.tex|.
 
We are gradually turning the source code for \LaTeX{} into a \LaTeX{}
document \emph{\sourcecode}.  This document includes an index of
\LaTeX{} commands and can be typeset from |source2e.tex|.
 
For more information about \TeX{} and \LaTeX, please contact your local
\TeX{} Users Group, or the international \TeX{} Users Group, P.~O.\
Box~869, Santa Barbara, CA 93102-0869, USA, Fax:~+1 805 963 8358,
E-mail:~tug@tug.org.
 
\section{Text fonts}
\label{Sec:text}
 
This section describes the commands available to class and package
writers for specifying and selecting fonts.
 
\subsection{Text font attributes}
 
Every text font in \LaTeX{} has five \emph{attributes}:
\begin{description}
 
\item[encoding] This specifies the order that characters appear in the
   font.  The two most common text encodings used in \LaTeX{} are
   Knuth's `\TeX{} text' encoding, and the `\TeX{} text extended'
   encoding developed by the \TeX{} Users Group members during a \TeX{}
   Conference at Cork in~1990 (hence its informal name `Cork encoding').
 
\item[family] The name for a collection of fonts, usually grouped under
   a common name by the font foundry.  For example, `Adobe Times', `ITC
   Garamond', and Knuth's `Computer Modern Roman' are all font families.
 
\item[series] How heavy or expanded a font is.  For example, `medium
   weight', `narrow' and `bold extended' are all series.
 
\item[shape] The form of the letters within a font family.  For
   example, `italic', `oblique' and `upright' (sometimes called
   `roman') are all font shapes.
 
\item[size] The design size of the font, for example `10pt'.
 
\end{description}
The possible values for these attributes are given short acronyms by
\LaTeX.  The most common values for the font encoding are:
\begin{center}
\begin{minipage}{.7\linewidth}
   \begin{tabular}{rl}
      |OT1|   & \TeX{} text  \\
      |T1|    & \TeX{} extended text \\
      |OML|   & \TeX{} math italic \\
      |OMS|   & \TeX{} math symbols \\
      |OMX|   & \TeX{} math large symbols \\
      |U|     & Unknown \\
      |L<xx>| & A local encoding
   \end{tabular}
\end{minipage}
\end{center}
The `local' encodings are intended for font encodings which are only
locally available, for example a font containing a university logo
in various sizes.
 
There are far too many font families to list them all, but some common
ones are:
\begin{center}
\begin{minipage}{.7\linewidth}
   \begin{tabular}{rl}
      |cmr|  & Computer Modern Roman \\
      |cmss| & Computer Modern Sans \\
      |cmtt| & Computer Modern Typewriter \\
      |cmm|  & Computer Modern Math Italic \\
      |cmsy| & Computer Modern Math Symbols \\
      |cmex| & Computer Modern Math Extensions \\
      |ptm|  & Adobe Times \\
      |phv|  & Adobe Helvetica \\
      |pcr|  & Adobe Courier
   \end{tabular}
\end{minipage}
\end{center}
The most common values for the font series are:
\begin{center}
\begin{minipage}{.7\linewidth}
   \begin{tabular}{rl}
      |m|    & Medium  \\
      |b|    & Bold  \\
      |bx|   & Bold extended \\
      |sb|   & Semi-bold \\
      |c|    & Condensed
   \end{tabular}
\end{minipage}
\end{center}
The most common values for the font shape are:
\begin{center}
\begin{minipage}{.7\linewidth}
   \begin{tabular}{rl}
      |n|    & Normal (that is `upright' or `roman') \\
      |it|   & Italic \\
      |sl|   & Slanted (or `oblique') \\
      |sc|   & Caps and small caps
   \end{tabular}
\end{minipage}
\end{center}
The font size is specified as a dimension, for example |10pt| or
|1.5in| or |3mm|.  These five parameters specify every \LaTeX{} font,
for example:
\begin{center}
   \begin{tabular}{@{}c@{\,}c@{\,}c@{\,}c@{\,}ccc@{}}
      \multicolumn{5}{@{}c}{\emph{\LaTeX{} specification}} &
      \emph{Font} &
      \emph{\TeX{} font name} \\
      |OT1| & |cmr| & |m| & |n| & |10pt| &
      Computer Modern Roman 10pt &
      |cmr10| \\
      |OT1| & |cmss| & |m| & |sl| & |12pt| &
      Computer Modern Sans Oblique 12pt &
      |cmssi12| \\
      |OML| & |cmm| & |m| & |it| & |10pt| &
      Computer Modern Math Italic 10pt &
      |cmmi10| \\
      |T1| & |ptm| & |b| & |it| & |18pt| &
      Adobe Times Bold Italic 18pt &
      |ptmbq at 18pt|
   \end{tabular}
\end{center}
These five parameters are displayed whenever \LaTeX{} gives an overfull
box warning, for example:
\begin{verbatim}
   Overfull \hbox (3.80855pt too wide) in paragraph at lines 314--318
   []\OT1/cmr/m/n/10 Normally [] and [] will be iden-ti-cal,
\end{verbatim}
The author commands for fonts set the five attributes:
\begin{center}
   \begin{tabular}{rcl}
      \emph{Author command} &
      \emph{Attribute} &
      \emph{Value in} |article| \emph{class} \\
      |\textrm{..}| or |\rmfamily|    & family & |cmr| \\
      |\textsf{..}| or |\sffamily|    & family & |cmss| \\
      |\texttt{..}| or |\ttfamily|    & family & |cmtt| \\
      |\textmd{..}| or |\mdseries|    & series & |m| \\
      |\textbf{..}| or |\bfseries|    & series & |bx| \\
      |\textup{..}| or |\upshape|     & shape  & |n| \\
      |\textit{..}| or |\itshape|     & shape  & |it| \\
      |\textsl{..}| or |\slshape|     & shape  & |sl| \\
      |\textsc{..}| or |\scshape|     & shape  & |sc| \\
      |\tiny|         & size   & |5pt| \\
      |\scriptsize|   & size   & |7pt| \\
      |\footnotesize| & size   & |8pt| \\
      |\small|        & size   & |9pt| \\
      |\normalsize|   & size   & |10pt| \\
      |\large|        & size   & |12pt| \\
      |\Large|        & size   & |14.4pt| \\
      |\LARGE|        & size   & |17.28pt| \\
      |\huge|         & size   & |20.74pt| \\
      |\Huge|         & size   & |24.88pt|
   \end{tabular}
\end{center}
The values used by these commands are determined by the document class,
using the parameters defined in Section~\ref{Sec:text.param}.
 
Note that there are no author commands for selecting new encodings.
These should be provided by packages, such as the |fontenc| package.
 
This section does not explain how \LaTeX{} font specifications are
turned into \TeX{} font names.  This is described in
Section~\ref{Sec:install}.
 
\subsection{Selection commands}
 
The low-level commands used to select a text font are as follows.
 
\begin{decl}
|\fontencoding| \arg{encoding} \\
|\fontfamily|  \arg{family}\\
|\fontseries| \arg{series}\\
|\fontshape| \arg{shape}\\
|\fontsize| \arg{size} \arg{baselineskip}
\end{decl}
 
Each of these commands sets one of the font
attributes; |\fontsize| also sets |\baselineskip|. The
actual font in use is not altered by these commands, but the current
attributes are used to determine which font to use after the next
|\selectfont| command.
 
\begin{decl}
|\selectfont|
\end{decl}
Selects a text font, based on the current values of the font attributes.
 
\emph{Warning}: there \emph{must} be a |\selectfont| command
immediately after any settings of the font parameters by (some of)
the five |\font<parameter>| commands, before any following text.
For example, it is legal to say:
\begin{verbatim}
   \fontfamily{ptm}\fontseries{b}\selectfont Some text.
\end{verbatim}
but it is \emph{not} legal to say:
\begin{verbatim}
   \fontfamily{ptm} Some \fontseries{b}\selectfont text.
\end{verbatim}
You may get unexpected results if you put text between a
|\font<parameter>| command and a |\selectfont|.
 
\begin{decl}
|\usefont| \arg{encoding} \arg{family} \arg{series} \arg{shape}
\end{decl}
A short hand for the equivalent |\font|\ldots{} commands followed by a
call to |\selectfont|.
 
 
\subsection{Internals}
 
The current values of the font attributes are held in internal macros.
 
\begin{decl}
      |\f@encoding| \\
      |\f@family|  \\
      |\f@series|  \\
      |\f@shape|  \\
      |\f@size|    \\
      |\f@baselineskip| \\
      |\tf@size|  \\
      |\sf@size|  \\
      |\ssf@size|
\end{decl}
 
These hold the current values of the encoding, the family, the series,
the shape, the size, the baseline skip, the main math size, the script
math size and the scriptscript math size. The last three are accessible
only within a formula; outside of math they may contain arbitrary
values.
 
For example, to set the size to 12 without
changing the baseline skip:
\begin{verbatim}
   \fontsize{12}{\f@baselineskip}
\end{verbatim}
However, you should \emph{never} alter the values of the internal
commands directly; they must only be modified using the low-level
commands like |\fontfamily|, |\fontseries|, etc. If you disobey this
warning you might produce code that loops.
 
\subsection{Parameters for author commands}
\label{Sec:text.param}
 
The parameter values set by author commands such as |\textrm| and
|\rmfamily|, etc.\ are not hard-wired into \LaTeX; instead these
commands use the values of a number of parameters set by the document
class and packages.  For example, |\rmdefault| is the name of the
default family selected by |\textrm| and |\rmfamily|. Thus to set a
document in Adobe Times, Helvetica and Courier, the document designer
specifies:
\begin{verbatim}
   \renewcommand{\rmdefault}{ptm}
   \renewcommand{\sfdefault}{phv}
   \renewcommand{\ttdefault}{pcr}
\end{verbatim}
 
\begin{decl}
 |\encodingdefault| \\
 |\familydefault|   \\
 |\seriesdefault|   \\
 |\shapedefault|
\end{decl}
The encoding, family, series and shape of the main body font.  By
default these are |OT1|, |\rmdefault|, |m| and |n|.  Note that since
the default family is |\rmdefault|, this means that changing
|\rmdefault| will change the main body font of the document.
 
\begin{decl}
 |\rmdefault|  \\
 |\sfdefault| \\
 |\ttdefault|
\end{decl}
The families selected by |\textrm|, |\rmfamily|, |\textsf|,
|\sffamily|, |\texttt| and |\ttfamily|.  By default these are |cmr|,
|cmss| and |cmtt|.
 
\begin{decl}
 |\bfdefault| \\
 |\mddefault|
\end{decl}
The series selected by |\textbf|, |\bfseries|, |\textmd| and
|\mdseries|.  By default these are |bx| and |m|.
These values are suitable for the default families used. If other
fonts are used as standard document fonts, for example, some
PostScript fonts, it might be necessary to adjust the value of
|\bfdefault| to |b| since only a few such families have `bold
extended' series.  An alternative (taken for the fonts provided by
|psnfss|) is to define silent substitutions from |bx| series to |b|
series with special |\DeclareFontShape| declarations and the |ssub|
size function, see Section~\ref{sec:sizefunct}.
 
 
\begin{decl}
 |\itdefault| \\
 |\sldefault| \\
 |\scdefault| \\
 |\updefault|
\end{decl}
The shapes selected by |\textit|, |\itshape|, |\textsl|, |\slshape|,
|\textsc|, |\scshape|, |\textup| and |\upshape|.  By default these are
|it|, |sl|, |sc| and |n|.
 
Note that there are no parameters for the size commands.  These should
be defined directly in class files, for example:
\begin{verbatim}
   \renewcommand{\normalsize}{\fontsize{10}{12}\selectfont}
\end{verbatim}
More elaborate examples (setting additional parameters when the text
size is changed) can be found in |classes.dtx| the source
documentation for the classes |article|, |report|, and |book|.
 
 
\subsection{Special font declaration commands}
 
\begin{decl}
|\DeclareFixedFont| \arg{cmd} \arg{ENC} \arg{family} \arg{series}
                    \arg{shape} \arg{size}
\end{decl}
 
Declares command \m{cmd} to be a font switch which selects the font
that is specified by the attributes \m{ENC}, \m{family}, \m{series},
\m{shape}, and \m{size}.
 
The font is selected without any adjustments to baselineskip and other
surrounding conditions.
 
This example makes |{\picturechar .}| select a small dot very quickly:
\begin{verbatim}
   \DeclareFixedFont{\picturechar}{OT1}{cmr}{m}{n}{5}
\end{verbatim}
 
 
\begin{decl}
|\DeclareTextFontCommand| \arg{cmd} \arg{font-switches}
\end{decl}
 
Declares command \m{cmd} to be a font command with one argument.
The current font attributes are locally modified by \m{font-switches}
and then the argument of \m{cmd} is typeset in the resulting new font.
 
Commands defined by |\DeclareTextFontCommand| automatically take care
of any necessary italic correction (on either side).
 
The following example shows how |\textrm| is defined by the kernel.
\begin{verbatim}
   \DeclareTextFontCommand{\textrm}{\rmfamily}
\end{verbatim}
 
To define a command that always typeset its argument in the italic
shape of the main document font you could declare:
\begin{verbatim}
   \DeclareTextFontCommand{\normalit}{\normalfont\itshape}
\end{verbatim}
 
This declaration can be used to change the meaning of a command; if
\m{cmd} is already defined, a log that it has been redefined
is put in the transcript file.
 
\begin{decl}
|\DeclareOldFontCommand| \arg{cmd} \arg{text-switch}
                                   \arg{math-switch}
\end{decl}
 
Declares command \m{cmd} to be a font switch (i.e.~used with the
syntax |{<cmd>...}|) having the definition \m{text-switch}
when used in text and the definition \m{math-switch} when used in a
formula.
Math alphabet commands, like |\mathit|, when used within \m{math-switch}
should not have an argument.  Their use in this argument causes their
semantics to change so that they here act as a font switch, as
required by the usage of the \m{cmd}.
 
This declaration is useful for setting up commands like |\rm| to behave
as they did in \LaTeX~2.09. We strongly urge you \emph{not} to misuse
this declaration to invent new font commands.
 
The following example defines |\it| to produce the italic shape of the
main document font if used in text and to switch to the font that would
normally be produced by the math alphabet |\mathit| if used in a
formula.
\begin{verbatim}
   \DeclareOldFontCommand{\it}{\normalfont\itshape}{\mathit}
\end{verbatim}
 
This declaration can be used to change the meaning of a command; if
\m{cmd} is already defined, a log that it has been redefined
is put in the transcript file.
 
 
\section{Math fonts}
\label{Sec:math}
 
This section describes the commands available to class and package
writers for specifying math fonts and math commands.
 
\subsection{Math font attributes}
 
The selection of fonts within math mode is quite different to that of
text fonts.
 
Some  math fonts are selected explicitly by one-argument commands such
as |\mathsf{max}| or |\mathbf{vec}|: these are called \emph{math
alphabets}.  These math alphabet commands affect only the font used
for letters and symbols of type |\mathalpha| (see
Section~\ref{Sec:math.commands}); other symbols within the argument
will be left unchanged.  The predefined math alphabets are:
\begin{center}
   \begin{tabular}{ccc}
      \emph{Alphabet} & \emph{Description} & \emph{Example} \\
      |\mathnormal|   & default      & $abcXYZ$ \\
      |\mathrm|       & roman        & $\mathrm{abcXYZ}$ \\
      |\mathbf|       & bold roman   & $\mathbf{abcXYZ}$ \\
      |\mathsf|       & sans serif   & $\mathsf{abcXYZ}$ \\
      |\mathit|       & text italic  & $\mathit{abcXYZ}$ \\
      |\mathtt|       & typewriter   & $\mathtt{abcXYZ}$ \\
      |\mathcal|      & calligraphic & $\mathcal{XYZ}$
   \end{tabular}
\end{center}
Other math fonts are selected implicitly by \TeX{} for symbols, with
commands such as |\oplus| (producing $\oplus$) or with straight
characters like |>>| or |+|.  Fonts containing such math symbols are
called \emph{math symbol fonts}.  The predefined math symbol fonts
are:
\begin{center}
   \begin{tabular}{ccc}
  \emph{Symbol font} & \emph{Description}         & \emph{Example} \\
      |operators|    & symbols from |\mathrm|     & $[\;+\;]$ \\
      |letters|      & symbols from |\mathnormal| & $<<\star>>$ \\
      |symbols|      & most \LaTeX{} symbols      & $\leq*\geq$ \\
      |largesymbols| & large symbols              & $\sum\prod\int$
   \end{tabular}
\end{center}
Some math fonts are both \emph{math alphabets} and \emph{math symbol
fonts}, for example |\mathrm| and |operators| are the same font, and
|\mathnormal| and |letters| are the same font.
 
Math fonts in \LaTeX{} have the same five attributes as text fonts:
encoding, family, series, shape and size.  However, there are no
commands that allow the attributes to be individually changed.
Instead,
the conversion from math fonts to these five attributes is controlled
by the \emph{math version}.  For example, the |normal| math version
maps:
\begin{center}
   \begin{tabular}{rlc@{ }c@{ }c@{ }c}
      \multicolumn{2}{c}{\emph{Math font}} &
      \multicolumn{4}{c}{\emph{External font}} \\
      \emph{Alphabets} & \emph{Symbol fonts} &
      \multicolumn{4}{c}{\emph{Attributes}} \\
      |\mathnormal| & |letters|      & |OML| & |cmm|  & |m|  & |it| \\
      |\mathrm|     & |operators|    & |OT1| & |cmr|  & |m|  & |n|  \\
      |\mathcal|    & |symbols|      & |OMS| & |cmsy| & |m|  & |n|  \\
                    & |largesymbols| & |OMX| & |cmex| & |m|  & |n|  \\
      |\mathbf|     &                & |OT1| & |cmr|  & |bx| & |n|  \\
      |\mathsf|     &                & |OT1| & |cmss| & |m|  & |n|  \\
      |\mathit|     &                & |OT1| & |cmr|  & |m|  & |it| \\
      |\mathtt|     &                & |OT1| & |cmtt| & |m|  & |n|
   \end{tabular}
\end{center}
The |bold| math version is similar except that it contains bold fonts.
The command |\boldmath| selects the |bold| math version.
 
Math versions can only be changed outside of math mode.
 
The two predefined math versions are:
\begin{center}
   \begin{tabular}{rl}
      |normal| & the default math version \\
      |bold|   & the bold math version
   \end{tabular}
\end{center}
Packages may define new math alphabets, math symbol fonts, and math
versions.  This section describes the commands for writing such
packages.
 
\subsection{Selection commands}
 
There are no commands for selecting symbol fonts.  Instead, these are
selected indirectly through symbol commands like |\oplus|.
Section~\ref{Sec:math.commands} explains how to define symbol commands.
 
\begin{decl}
      |\mathnormal{<math>}|\\
      |\mathcal{<math>}|\\
      |\mathrm{<math>}|\\
      |\mathbf{<math>}|\\
      |\mathsf{<math>}|\\
      |\mathit{<math>}|\\
      |\mathtt{<math>}|
\end{decl}
Each math alphabet is a command which can only be used inside math
mode.  For example, |$x + \mathsf{y} + \mathcal{Z}$| produces $x +
\mathsf{y} + \mathcal{Z}$.
 
\begin{decl}
|\mathversion{<version-name>}|
\end{decl}
This command selects a math version; it can only be used outside math
mode.  For example, |\boldmath| is defined to be |\mathversion{bold}|.
 
\subsection{Declaring math versions}
 
\begin{decl}
|\DeclareMathVersion| \arg{version-name}
\end{decl}
 
Defines \m{version-name} to be a math version.
 
The newly declared version is initialised with the defaults for all
symbol fonts and math alphabets
declared so far (see the commands
|\DeclareSymbolFont| and |\DeclareMathAlphabet|).
 
If used on an already existing version, an information message is
written to the transcript file and all previous |\SetSymbolFont| or
|\SetMathAlphabet| declarations for this version are overwritten by
the math alphabet and symbol font defaults, i.e.~one ends up with a
virgin math version.
 
Example:
\begin{verbatim}
   \DeclareMathVersion{normal}
\end{verbatim}
 
 
\subsection{Declaring math alphabets}
 
\begin{decl}
|\DeclareMathAlphabet| \arg{math-alph} \arg{encoding} \arg{family}
                       \arg{series} \arg{shape}
\end{decl}
 
Defines \m{math-alph} to be a new math alphabet.
 
The arguments \m{encoding} \m{family} \m{series} \m{shape} are the
default values for this math alphabet in all math versions; these can
be reset later for a particular math version by a |\SetMathAlphabet|
command.  If \m{shape} is empty then the \m{math-alph} is declared to
be invalid in all versions, unless it is set by a later
|\SetMathAlphabet| command.
 
Checks that \m{math-alph} can be used and that \m{encoding} is a valid
encoding scheme.
 
In these examples, |\foo| is defined everywhere but |\baz|,
by default, is defined nowhere.
\begin{verbatim}
   \DeclareMathAlphabet{\foo}{OT1}{cmtt}{m}{n}
   \DeclareMathAlphabet{\baz}{OT1}{}{}{}
\end{verbatim}
 
 
\begin{decl}
|\SetMathAlphabet| \arg{math-alph} \arg{version-name}\\
        \null\hfill\arg{encoding} \arg{family} \arg{series} \arg{shape}
\end{decl}
 
Changes, or sets, the font for the math alphabet \m{math-alph} in math
version \m{version-name} to \m{encoding}\m{family}\m{series}\m{shape}.
 
Checks that \m{math-alph} is a math alphabet, \m{version-name} is a
math version and \m{encoding} is a known encoding scheme.
 
This example defines |\baz| for the `normal' math version only:
\begin{verbatim}
   \SetMathAlphabet\baz{normal}{OT1}{cmss}{m}{n}
\end{verbatim}
 
Note that this declaration is not used for all math alphabets:
Section~\ref{sec:symalph} describes |\DeclareSymbolFontAlphabet|,
which is used to set up math alphabets contained in fonts which have
are declared as symbol fonts.
 
 
\subsection{Declaring symbol fonts}
\label{sec:symalph}
 
\begin{decl}
|\DeclareSymbolFont| \arg{sym-font-name} \arg{encoding} \arg{family}
                     \arg{series} \arg{shape}
\end{decl}
 
Defines \m{sym-font-name} to be a new symbol font.
 
The arguments \m{encoding} \m{family} \m{series} \m{shape} are the
default values for this symbol font in all math versions; these can be
reset later for a particular math version by a |\SetSymbolFont|
command.
 
Checks that \m{sym-font-name} can be used and that \m{encoding} is a
declared encoding scheme.
 
 
For example, the following sets up the first four standard math symbol
fonts:
\begin{verbatim}
   \DeclareSymbolFont{operators}{OT1}{cmr}{m}{n}
   \DeclareSymbolFont{letters}{OML}{cmm}{m}{it}
   \DeclareSymbolFont{symbols}{OMS}{cmsy}{m}{n}
   \DeclareSymbolFont{largesymbols}{OMX}{cmex}{m}{n}
\end{verbatim}
 
 
\begin{decl}
|\SetSymbolFont| \arg{sym-font-name} \arg{version name}\\
     \null\hfill \arg{encoding} \arg{family} \arg{series} \arg{shape}
\end{decl}
 
Changes the symbol font \m{sym-font-name} for math version \m{version
  name} to \m{encoding} \m{family} \m{series} \m{shape}.
 
Checks that \m{sym-font-name} is a symbol font, \m{version name} is a
known math version and \m{encoding} is a declared encoding scheme.
 
For example, the following come from the set up of the `bold'
math version:
\begin{verbatim}
   \SetSymbolFont{operators}{bold}{OT1}{cmr}{bx}{n}
   \SetSymbolFont{letters}{bold}{OML}{cmm}{b}{it}
\end{verbatim}
 
 
\begin{decl}
|\DeclareSymbolFontAlphabet| \arg{math-alph} \arg{sym-font-name}
\end{decl}
 
Allows the previously declared symbol font \m{sym-font-name} to be
also the math alphabet \m{id} (in all math versions).
 
This declaration should be used in preference to
|\DeclareMathAlphabet| and |\SetMathAlphabet| when a math alphabet is
the same as a symbol font; this is because it makes better use of the
limited number (only 16) of \TeX's math groups.
 
Checks that \m{math-alph} can be defined and that
\m{sym-font-name} is a symbol font.
 
Example:
\begin{verbatim}
   \DeclareSymbolFontAlphabet{\mathrm}{operators}
   \DeclareSymbolFontAlphabet{\mathcal}{symbols}
\end{verbatim}
 
 
\subsection{Declaring math symbols}
\label{Sec:math.commands}
 
\begin{decl}
|\DeclareMathSymbol| \arg{symbol} \arg{type} \arg{sym-font-name}
                     \arg{slot}
\end{decl}
 
The \m{symbol} can be either a single character such as `|>>|', or a
macro name, such as |\sum|.
 
Defines the \m{symbol} to be a math symbol of type
\m{type} in slot \m{slot} of symbol font \m{sym-font-name}. The
\m{type} can be given as a number or as a command:
\begin{center}
   \begin{tabular}{ccc}
      \emph{Type} & \emph{Meaning} & \emph{Example} \\
      |0| or |\mathord  | & Ordinary           & $\alpha$ \\
      |1| or |\mathop   | & Large operator     & $\sum$ \\
      |2| or |\mathbin  | & Binary operation   & $\times$ \\
      |3| or |\mathrel  | & Relation           & $\leq$ \\
      |4| or |\mathopen | & Opening            & $\langle$ \\
      |5| or |\mathclose| & Closing            & $\rangle$ \\
      |6| or |\mathpunct| & Punctuation        & $;$ \\
      |7| or |\mathalpha| & Alphabet character & $A$
   \end{tabular}
\end{center}
Only symbols of type |\mathalpha| will be affected by math alphabet
commands: within the argument of a math alphabet command they will
produce the character in slot \m{slot} of that math alphabet's font.
Symbols of other types will always produce the same symbol
(within one math version).
 
|\DeclareMathSymbol| allows a macro \m{symbol} to be redefined only if
it was previously defined to be a math symbol.
It also checks that the \m{sym-font-name} is a
declared symbol font.
 
Example:
\begin{verbatim}
   \DeclareMathSymbol{\alpha}{0}{letters}{"0B}
   \DeclareMathSymbol{\lessdot}{\mathbin}{AMSb}{"0C}
   \DeclareMathSymbol{\foo}{\mathalpha}{AMSb}{"0C}
\end{verbatim}
 
 
\begin{decl}
 |\DeclareMathDelimiter| \arg{cmd}   \arg{type}
               \arg{sym-font-name-1} \arg{slot-1}\\
  \null\hfill  \arg{sym-font-name-2} \arg{slot-2}
\end{decl}
Defines \m{cmd} to be a math delimiter where the small
variant is in slot \m{slot-1} of symbol font
\m{sym-font-name-1} and the large variant is in slot
\m{slot-2} of symbol font \m{sym-font-name-2}.  Both symbol fonts
must have been declared previously.
 
Checks that \m{sym-font-name-i} are both
declared symbol fonts.
 
If \TeX{} is not looking for a delimiter, \m{cmd} is treated just as
if it had been defined with |\DeclareMathSymbol| using
\m{type},\m{sym-font-name-1} and \m{slot-1}.  In other words, if a
command is defined as a delimiter then this automatically defines it
as a math symbol.
 
Example:
\begin{verbatim}
   \DeclareMathDelimiter{\langle}{\mathopen}{symbols}{"68}
                                            {largesymbols}{"0A}
\end{verbatim}
 
 
\begin{decl}
|\DeclareMathAccent| \arg{cmd} \arg{type} \arg{sym-font-name} \arg{slot}
\end{decl}
 
Defines \m{cmd} to act as a math accent.
 
The accent character comes from slot \m{slot} in
\m{sym-font-name}. The \m{type} can be either
|\mathord| or |\mathalpha|; in the latter case the accent character
changes font when used in a math alphabet.
 
Example:
\begin{verbatim}
   \DeclareMathAccent{\acute}{\mathalpha}{operators}{"13}
   \DeclareMathAccent{\vec}{\mathord}{letters}{"7E}
\end{verbatim}
 
 
\begin{decl}
|\DeclareMathRadical| \arg{cmd}
                    \arg{sym-font-name-1} \arg{slot-1}\\
       \null\hfill  \arg{sym-font-name-2} \arg{slot-2}
\end{decl}
 
Defines \m{cmd} to be a radical where the small variant is in
slot \m{slot-1} of symbol font \m{sym-font-name-1} and the
large variant is in slot \m{slot-2} of symbol font
\m{sym-font-name-2}.  Both symbol fonts must have been declared
previously.
 
Example (probably the only use for it!):
\begin{verbatim}
   \DeclareMathRadical{\sqrt}{symbols}{"70}{largesymbols}{"70}
\end{verbatim}
 
\subsection{Declaring math sizes}
 
\begin{decl}
|\DeclareMathSizes| \arg{t-size} \arg{mt-size} \arg{s-size}
                    \arg{ss-size}
\end{decl}
 
Declares that \m{mt-size} is the math text size, \m{s-size} is the
script size and \m{ss-size} the scriptscript size to be used in math,
when \m{t-size} is the current text size. For text sizes for which no
such declaration is given the script and scriptscript size will be
calculated and then fonts are loaded for the calculated sizes or the
best approximation (this may result in a warning message).
 
Normally, \m{t-size} and \m{mt-size} will be identical; however, if,
for example, PostScript text fonts are mixed with bit-map math fonts
then you may not have available a \m{mt-size} for every \m{t-size}.
 
Example:
\begin{verbatim}
   \DeclareMathSizes{13.82}{12.4}{10}{7}
\end{verbatim}
 
\section{Font installation}
\label{Sec:install}
 
This section explains how \LaTeX's font attributes are turned into
\TeX{} font specifications.
 
\subsection{Font definition files}
 
The description of how \LaTeX{} font attributes are turned into \TeX{}
fonts is usually kept in a \emph{font definition} (|.fd|) file.  The
|.fd| file for family \m{family} in encoding \m{encoding} is called
|<ENC><family>.fd|, for example |OT1cmr.fd| for Computer Modern Roman
or |T1ptm.fd| for Adobe Times.
 
Whenever \LaTeX{} encounters an encoding/family combination that it
does not know (e.g.~if the document designer says
|\fontfamily{ptm}\selectfont|) then \LaTeX{} attempts to load the
appropriate |.fd| file.  ``Not known'' means: there was no
|\DeclareFontFamily| declaration issued for this encoding/family
combination.
 
The declarations in the |.fd| file are responsible for telling
\LaTeX{} how to load fonts for that encoding/family combination.  If
the |.fd| file could not be found, a warning is issued and font
substitutions are made.
 
 
\subsection{Font definition file commands}
 
The |.fd| files should contain only commands from this section.
Note that those commands can also be used outside a |.fd| file:
they can be put in package or class files, or even in the preamble
of a document.
 
\begin{decl}
   |\ProvidesFile{<file-name>}[<release-info>]|
\end{decl}
The |.fd| file should announce itself with a |\ProvidesFile| command,
as described in \emph{\clsguide}.  For example:
\begin{verbatim}
   \ProvidesFile{T1ptm.fd}[1994/06/01 Adobe Times font definitions]
\end{verbatim}
 
\begin{decl}
|\DeclareFontFamily| \arg{encoding} \arg{family} \arg{loading-settings}
\end{decl}
 
Declares a font family \m{family} to be available in encoding scheme
\m{encoding}.
 
The \m{loading-settings} are executed immediately after loading any
font with this encoding and family.
 
Checks that \m{encoding} was previously declared.
 
This example refers to the Computer Modern Typewriter font family in
the Cork encoding:
\begin{verbatim}
   \DeclareFontFamily{T1}{cmtt}{\hyphenchar\font=-1}
\end{verbatim}
 
Each |.fd| file should contain exactly one |\DeclareFontFamily|
command, and it should be for the appropriate encoding/family
combination.
 
\begin{decl}
|\DeclareFontShape| \arg{encoding} \arg{family} \arg{series}
                    \arg{shape}\\
        \null\hfill \arg{loading-info} \arg{loading-option}
\end{decl}
 
Declares a font shape combination; here \m{loading-info} contains the
information that combines sizes with external fonts. The syntax is
complex and is described in Section~\ref{sec:loadinfo} below.
 
The \m{loading-settings} are executed after loading any font with this
font shape.  They are executed immediately after the
`loading-settings' which were declared by |\DeclareFontFamily| and so
they can be used to overwrite the settings made at the family level.
 
Checks that the combination \m{encoding}\m{family} was previously
declared via |\DeclareFontFamily|.
 
Example:
\begin{verbatim}
   \DeclareFontShape{OT1}{cmr}{m}{sl}{%
             <<5-8>> sub * cmr/m/n
             <<8>> cmsl8
             <<9>> cmsl9
             <<10>> <<10.95>> cmsl10
             <<12>> <<14.4>> <<17.28>> <<20.74>> <<24.88>> cmsl12
             }{}
\end{verbatim}
An |.fd| file can contain any number of |\DeclareFontShape| commands,
which should be for the appropriate \m{encoding} and \m{family}.
 
\subsection{Font file loading information}
\label{sec:loadinfo}
 
The information which tells \LaTeX{} exactly which font ({\tt .tfm})
files to load is contained in the \m{loading-info} part of a
|\DeclareFontShape| declaration. This part consists of one or more
\m{fontshape-decl}s, each of which has the following form:
 
\begin{center}
\begin{tabular}{r@{ $::=$ }l}
 \m{fontshape-decl} &  \m{size-infos} \m{font-info} \\
 \m{size-infos}     &  \m{size-infos} \m{size-info} $\mid$
                       \m{size-info} \\
 \m{size-info}      & ``|<<|''  \m{number-or-range} ``|>>|'' \\
 \m{font-info}      & $[$ \m{size-function} ``|*|''  $]$
                      $[$ ``|[|'' \m{optarg} ``|]|'' $]$ \m{fontarg} \\
\end{tabular}
\end{center}
The \m{number-or-range} denotes the size or size-range
for which this entry applies.
 
If it contains a hyphen it is a
range: lower bound on the left (if missing, zero implied), upper bound
on the right (if missing, $\infty$ implied). For ranges, the
upper bound is \emph{not} included in the range and the lower bound is.
 
Examples:
\begin{center}
\begin{tabular}{lll}
   |<<10>>|     &  simple size& 10pt only\\
   |<<-8>>|     &  range& all sizes less than 8pt\\
   |<<8-14.4>>| &  range& all sizes greater than or equal to 8pt\\
              &       & \ but less than 14.4pt\\
   |<<14.4->>|  &  range& all sizes greater than or equal 14.4pt
\end{tabular}
\end{center}
If more than one \m{size-info} entry follows without any
intervening \m{font-info}, they all share the next \m{font-info}.
 
The \m{size-function}, if present, handles the use of \m{font-info}.
If not present, the `empty' \m{size-function} is assumed.
 
All the \m{size-info}s are inspected in the order in which they appear
in the font shape declaration. If a \m{size-info} matches the
requested size, its \m{size-function} is executed. If |\external@font|
is non-empty afterwards this process stops, otherwise the next
\m{size-info} is inspected. (See also |\DeclareSizeFunction|.)
 
If this process does not lead to a non-empty |\external@font|,
\LaTeX{} tries the nearest simple size. If the entry contains only
ranges an error is returned.
 
 
\subsection{Size functions}
\label{sec:sizefunct}
 
 
\LaTeX{} provides the following size functions, whose `inputs' are
\m{fontarg} and \m{optarg} (when present).
 
\begin{description}
\item[`' (empty)]
Load the external font \m{fontarg} at the user-requested size. If
\m{optarg} is present, it is used as the scale-factor.
 
\item[s]
Like the empty function but without terminal warnings, only
loggings.
 
\item[gen]
Generates the external font from \m{fontarg} followed by
the user-requested size, e.g. |<<8>> <<9>> <<10>> gen * cmtt|
 
\item[sgen]
Like the `gen' function but without terminal warnings, only loggings.
 
\item[sub]
Tries to load a font from a different font shape declaration given by
\m{fontarg} in the form \m{family}|/|\m{series}|/|\m{shape}.
 
\item[ssub]
Silent variant of `sub', only loggings.
 
\item[subf]
Like the empty function but issues a warning that it has to substitute
the external font \m{fontarg} because the desired font shape was not
available in the requested size.
 
\item[ssubf]
Silent variant of `subf', only loggings.
 
\item[fixed]
Load font \m{fontarg} as is, disregarding the user-requested size.
If present, \m{optarg} gives the ``at \ldots pt'' size to be used.
 
\item[sfixed]
Silent variant of `fixed', only loggings.
 
\end{description}
 
Examples for the use of most of the above size functions can be found
in the file |cmfonts.fdd|---the source for the standard |.fd| files
describing the Computer Modern fonts by Donald Knuth.
 
 
\begin{decl}
|\DeclareSizeFunction| \arg{name} \arg{code}
\end{decl}
 
Declares a size-function \m{name} for use in |\DeclareFontShape|
commands.  The interface is still under development but there should be
no real need to a define new size functions.
 
The \m{code} is executed when the size or size-range in
|\DeclareFontShape| matches the user-requested size.
 
The arguments of the size-function are automatically parsed and placed
into |\mandatory@arg| and |\optional@arg| for use in \m{code}. Also
available, of course, is |\f@size|, which is the user-requested size.
 
To signal success \m{code} must define the command |\external@font|
to contain the external name and any scaling options (if present) for
the font to be loaded.
 
This example sets up the `empty' size function (simplified):
\begin{verbatim}
   \DeclareSizeFunction{}
           {\edef\external@font{\mandatory@arg\space at\f@size}
\end{verbatim}
 
 
\section{Encodings}
\label{Sec:encode}
 
This section explains how to declare and use new font encodings and
how to declare commands for use with particular encodings.
 
\subsection{The \textsf{fontenc} package}
 
Users can select new font encodings using the |fontenc| package.  The
|fontenc| package has options for encodings; the last option becomes
the default encoding.  For example, to use the |OT2| (Washington
University Cyrillic encoding) and |T1| encodings, with |T1| as the
default, an author types:
\begin{verbatim}
   \usepackage[OT2,T1]{fontenc}
\end{verbatim}
 
This package loads the \emph{encoding definition files}
(|<ENC>enc.def| files) for each font encoding \m{ENC} given as an
option but not already declared; it also sets |\encodingdefault| to be
the last encoding in the option list.
 
\LaTeX{} currently predefines the |OT1| and |T1| text encodings, and
provides the files |OT1enc.def| and |T1enc.def|. Other encoding
set-ups might be added to the distribution at a later stage.
 
Thus the example above loads the file |OT2enc.def| and sets
|\encodingdefault| to |T1|.
 
\subsection{Encoding definition files}
\label{Sec:encode.def}
 
{\bf Warning:} Some aspects of the contents of font definition files
are still under development.  Therefore, the current versions of the
files |OT1enc.def| and |T1enc.def| are temporary versions and should
not be used as models for producing further such files.  For further
information you should read the documentation in |ltoutenc.dtx|.
 
The declarations in the encoding definition file are  responsible for
declaring the encoding and telling \LaTeX{} how to produce characters
in this encoding.
 
The |<ENC>enc.def| files should contain only commands from this
section.  As with the font definition file commands, it is also
possible (although normally not necessary) to use these declarations
directly within a class or package file.
 
\begin{decl}
   |\ProvidesFile{<file-name>}[<release-info>]|
\end{decl}
The |<ENC>enc.def| file should announce itself with a |\ProvidesFile|
command, described in \emph{\clsguide}.  For example:
\begin{verbatim}
   \ProvidesFile{OT2enc.def}
                [1994/06/01 Washington Univ. Cyrillic encoding]
\end{verbatim}
 
 
\begin{decl}
|\DeclareFontEncoding| \arg{encoding} \arg{text-settings}
                       \arg{math-settings}
\end{decl}
 
Declares a new encoding scheme \m{encoding}.
 
The \m{text-settings} are
declarations which are executed every time |\selectfont| changes the
encoding to be \m{encoding}.
 
The \m{math-settings} are similar but are for math alphabets. They
are executed whenever a math alphabet with this encoding is
called.
 
Spaces within the arguments are ignored to avoid surplus
spaces in the document. If a real space is necessary
use |\space|.
 
Example:
\begin{verbatim}
   \DeclareFontEncoding{OT1}{}{}
\end{verbatim}
 
Some author commands need to change their definition depending on
which encoding is currently in use.  For example, in the |OT1|
encoding, the letter `\AE' is in slot |"1D|, whereas in the |T1|
encoding it is in slot |"C6|.  So the definition of |\AE| has to
change depending on whether the current encoding is |OT1| or |T1|.
The following commands allow this to happen.
 
\begin{decl}
   |\DeclareTextCommand| \arg{cmd} \arg{encoding}
   \oarg{num} \oarg{default} \arg{definition}
\end{decl}
This command is like |\newcommand|, except that it defines a
command which is specific to one encoding.  For example, the
definition of |\aa| in the |OT1| encoding is:
\begin{verbatim}
  \DeclareTextCommand{\aa}{OT1}{{\accent23a}}
\end{verbatim}
|\DeclareTextCommands| takes the same optional arguments as
|\newcommand|.
 
The resulting command is
robust, even if the code in \m{definition} is fragile.
 
It does not produce an error if the command has already
been defined but logs the redefinition in the transcript file.
 
\begin{decl}[1994/12/01]
   |\ProvideTextCommand| \arg{cmd} \arg{encoding}
   \oarg{num} \oarg{default} \arg{definition}
\end{decl}
This command is the same as |\DeclareTextCommand|, except that
if \m{cmd} is already defined in encoding \m{encoding}, then the
definition is ignored.
 
\begin{decl}
   |\DeclareTextSymbol| \arg{cmd} \arg{encoding} \arg{slot}
\end{decl}
This command defines a text symbol with slot \m{slot} in the
encoding.  For example, the definition of |\ss| in the |OT1| encoding
is:
\begin{verbatim}
   \DeclareTextSymbol{\ss}{OT1}{25}
\end{verbatim}
It does not produce an error if the command has already
been defined but logs the redefinition in the transcript file.
 
\begin{decl}
   |\DeclareTextAccent| \arg{cmd} \arg{encoding} \arg{slot}
\end{decl}
This command declares a text accent, with the accent taken from slot
\m{slot} in the encoding.  For example, the definition of |\"| in
the |OT1| encoding is:
\begin{verbatim}
   \DeclareTextAccent{\"}{OT1}{127}
\end{verbatim}
It does not produce an error if the command has already
been defined but logs the redefinition in the transcript file.
 
\begin{decl}
   |\DeclareTextComposite| \arg{cmd} \arg{encoding} \arg{letter}
                           \arg{slot}
\end{decl}
This command declares that the composite letter formed from applying
\m{cmd} to \m{letter} is defined to be simply slot \m{slot} in the
encoding.   The \m{letter} should be a single letter (such as |a|) or
a single command (such as |\i|).
 
 
For example, the definition of |\'{a}|
in the |T1| encoding could be declared like this:
\begin{verbatim}
   \DeclareTextComposite{\'}{T1}{a}{225}
\end{verbatim}
 
The \m{cmd} should have been previously declared with
|\DeclareTextAccent|, or as a one-argument |\DeclareTextCommand|.
 
\begin{decl}[1994/12/01]
   |\DeclareTextCompositeCommand| \arg{cmd} \arg{encoding} \arg{letter}
                           \arg{definition}
\end{decl}
This is a more general form of |\DeclareTextComposite|, which allows for 
an arbitrary \m{definition}, not just a \m{slot}.  The main use for this 
is to allow accents on |i| to act like accents on |\i|, for example:
\begin{verbatim}
   \DeclareTextCompositeCommand{\'}{OT1}{i}{\'\i}
\end{verbatim}
It has the same restrictions as |\DeclareTextComposite|.
 
\subsection{Default definitions}
 
\NEWdescription{1994/12/01}
   The commands in \m{ENC}|enc.def| files allow encoding-specific 
   commands to be defined, but they do not allow commands to be used in 
   other encodings.  For example, the OMS encoding contains the symbol 
   `\S', but we need to be able to use the command |\S| in any encoding, 
   not just OMS.
   
   To allow this, \LaTeX{} has commands for giving default definitions 
   for commands, which are used when the command is not defined in the 
   current encoding.  For example, the default encoding for |\S| is OMS, 
   and so in an encoding (such as OT1) which does not contain |\S|, the 
   OMS version is selected.  But in an encoding (such as T1) which does 
   contain |\S|, the version for that encoding is used.
   
   \emph{Note} these commands should \emph{not} occur in 
   \m{ENC}|enc.def| files, since these should only define commands 
   for that encoding.  They should instead be placed in packages.
 
\begin{decl}[1994/12/01]
   |\DeclareTextCommandDefault| \arg{cmd} \arg{definition} \\
\end{decl}
This command allows an encoding-specific command to be given a default 
definition.  For example, the default definition for |\copyright| is 
defined be be a circled `c' with:
\begin{verbatim}
   \DeclareTextCommandDefault{\copyright}{\textcircled{c}}
\end{verbatim}
\begin{decl}[1994/12/01]
   |\DeclareTextAccentDefault| \arg{cmd} \arg{encoding} \\
   |\DeclareTextSymbolDefault| \arg{cmd} \arg{encoding}
\end{decl}
These commands allow an encoding-specific command to be given a
default encoding.  For example, the default encoding for |\"| and
|\ae| is set to be OT1 by:
\begin{verbatim}
   \DeclareTextAccentDefault{\"}{OT1}
   \DeclareTextSymbolDefault{\ae}{OT1}
\end{verbatim}
Note that |\DeclareTextAccentDefault| can be used on any one-argument 
encoding-specific command, not just those defined with 
|\DeclareTextAccent|.  Similarly, |\DeclareTextSymbolDefault| can be used 
on any encoding-specific command with no arguments, not just those 
defined with |\DeclareTextSymbol|.
 
For more examples of these definitions, see |ltoutenc.dtx|.  
 
\begin{decl}[1994/12/01]
   |\ProvideTextCommandDefault| \arg{cmd} \arg{definition}
\end{decl}
This command is the same as |\DeclareTextCommandDefault|, except that if 
the command already has a default definition, then the definition is 
ignored.  This is useful to give `faked' definitions of symbols which may 
be given `real' definitions by other packages.  For example, a package 
might give a fake definition of |\textonequarter| by saying:
\begin{verbatim}
   \ProvideTextCommandDefault{\textonequarter}{$\m@th\frac14$}
\end{verbatim}
 
 
\subsection{Encoding defaults}
 
\begin{decl}
|\DeclareFontEncodingDefaults| \arg{text-settings} \arg{math-settings}
\end{decl}
 
Declares \m{text-settings} and \m{math-settings} for all encoding
schemes.  These are executed before the encoding scheme dependent ones
are executed so that one can use the defaults for the major cases and
overwrite them if necessary using |\DeclareFontEncoding|.
 
If |\relax| is used as an argument, the current setting of this default
is left unchanged.
 
This example is used by amsfonts.sty for accent positioning; it changes
only the math settings:
\begin{verbatim}
   \DeclareFontEncodingDefaults{\relax}{\def\accentclass@{7}}
\end{verbatim}
 
 
\begin{decl}
|\DeclareFontSubstitution|  \arg{encoding} \arg{family} \arg{series}
                            \arg{shape}
\end{decl}
 
Declares the default values for font substitution which will be used
when a font with encoding \m{encoding}
should be loaded but no font can be found with the current
attributes.
 
These substitutions are local to the encoding scheme because the
encoding scheme is never substituted!  They are tried in the order
\m{shape} then  \m{series} and finally \m{family}.
 
If no defaults are set up for an encoding, the values given by
|\DeclareErrorFont| are used.
 
The font specification for
\m{encoding}\m{family}\m{series}\m{shape}
must have been defined by |\DeclareFontShape|
before the |\begin{document}| is reached.
 
Example:
\begin{verbatim}
   \DeclareFontSubstitution{T1}{cmr}{m}{n}
\end{verbatim}
 
 
\section{Miscellania}
\label{Sec:misc}
 
 
This section covers the remaining font commands of \LaTeX.
 
\subsection{Font substitution}
 
\begin{decl}
|\DeclareErrorFont| \arg{encoding} \arg{family} \arg{series}
                    \arg{shape} \arg{size}
\end{decl}
 
Declares \m{encoding}\m{family}\m{series}\m{shape} to be the font shape
used in cases where the standard substitution mechanism fails
(i.e.~would loop). For the standard mechanism see the command
|\DeclareFontSubstitution| above.
 
The font specification for
\m{encoding}\m{family}\m{series}\m{shape}
must have been defined by |\DeclareFontShape|
before the |\begin{document}| is reached.
 
Example:
\begin{verbatim}
   \DeclareErrorFont{OT1}{cmr}{m}{n}{10}
\end{verbatim}
 
 
\begin{decl}
|\fontsubfuzz|
\end{decl}
 
This parameter is used to decide whether or not to produce a terminal
warning if a font size substitution takes place. If the difference
between the requested and the chosen size is less than |\fontsubfuzz|
the warning is only written to the transcript file. The default value
is |0.4pt|.  This can be redefined with |\renewcommand|, for example:
\begin{verbatim}
   \renewcommand{\fontsubfuzz}{0pt}   % always warn
\end{verbatim}
 
\subsection{Preloading}
 
\begin{decl}
|\DeclarePreloadSizes| \arg{encoding} \arg{family} \arg{series}
                       \arg{shape}
\arg{size-list}
\end{decl}
 
 
Specifies the fonts that should be preloaded by the format.  These
commands should be put in a |preload.cfg| file, which is read in when
the \LaTeX{} format is being built.
Read |preload.dtx| for more information on how to built such a
configuration file.
 
Example:
\begin{verbatim}
   \DeclarePreloadSizes{OT1}{cmr}{m}{sl}{10,10.95,12}
\end{verbatim}
 
 
\subsection{Naming conventions}
 
\begin{itemize}
\item
Math alphabet commands all start with |\math...|: examples are
|\mathbf|, |\mathcal|, etc.
 
\item
The text font changing commands with arguments all start with
|\text...|: e.g.~|\textbf| and |\textrm|.  The exception to this is
|\emph|, since it occurs very commonly in author documents and so
deserves a shorter name.
 
\item
Names for encoding schemes are strings of up to three letters, all
upper case.  The \LaTeX3 project reserves the use of encodings
starting with |T| (standard 256-long text encodings), |M| (standard
256-long math encodings), |S| (standard 256-long symbol encodings),
|OT| (standard 128-long text encodings) and |OM| (standard 128-long
math encodings).  Please do not use the above starting letters for
non-portable encodings.  If new standard encoding emerge then we
shall add them in a later release of \LaTeX.
 
Encoding schemes which are local to a site should start with |L|.
 
\item
Font family names should contain up to five lower case letters.
Where possible, these should conform to the \emph{Filenames for fonts}
font naming scheme.
 
\item
Font series names should contain up to four lower case letters.
 
\item
Font shapes should contain up to two letters lower case.
 
\item
Names for symbol fonts are built from lower and upper case letters
with no restriction.
\end{itemize}
 
Whenever possible, you should use the series and shape names suggested
in \emph{\LaTeXcomp} since this will make it easier to combine new
fonts with existing fonts.
 
\NEWdescription{1994/12/01}
   Where possible, text symbols should be named as |\text| followed by
   the Adobe glyph name: for example |\textonequarter| or 
   |\textsterling|.  Similarly, math symbols should be named as 
   |\math| followed by the glyph name, for example |\mathonequarter| 
   or |\mathsterling|.  Commands which can be used in text or math can
   then be defined using |\ifmmode|, for example:
\begin{verbatim}
   \DeclareRobustCommand{\pounds}{%
      \ifmmode \mathsterling \else \textsterling \fi
   }
\end{verbatim}
   Note that commands defined in this way must be robust, in case they
   get put into a section title or other moving argument.

   
\begin{thebibliography}{1}
 
\bibitem{A-W:GMS94}
Michel Goossens, Frank Mittelbach and Alexander Samarin.
\newblock {\em The {\LaTeX} Companion}.
\newblock Addison-Wesley, Reading, Massachusetts, 1994.
 
\bibitem{tub:DKn89}
Donald~E. Knuth.
\newblock Typesetting concrete mathematics.
\newblock {\em {TUG}boat}, 10(1):31--36, April 1989.
 
\bibitem{A-W:LLa94}
Leslie Lamport.
\newblock {\em {\LaTeX:} A Document Preparation System}.
\newblock Addison-Wesley, Reading, Massachusetts, second edition, 1994.
 
\end{thebibliography}
 
\end{document}
 
