% -*-latex-*-
% Document name: /home2/boris/tla/Tex/mm2tex/mmstatus.tex
% Creator: Terry L Anderson 201-580-4428 [tla@seldon]
% Creation Date: Mon Jun  3 14:24:11 1991
\documentstyle[]{tm}
\newcommand{\X}[1]{{#1}\index{{#1}}}
\title{MM to \LaTeX\ Translator}
\input authors.sty
\input bitmap.sty
\input mymacros.sty
\authordata{tla}
%\chargecase{978899-0100}
%\filecase{61093}
%\documentno{900000}{02}{TMS}
\keywords{\TeX , \LaTeX , mm, troff}
\mercurycode{CMP}
\date{\today}
\memotype{TECHNICAL MEMORANDUM}
\abstract{Mm2tex is a filter to convert documents written in troff
using the mm macro package into \LaTeX\ using the tm.sty style package.
}
\organizationalapproval
%\markdraft*
\markproprietary
\begin{document}
\bibliographystyle{atttm}
\makehead
\makeabstract
\section{Introduction}
The {\bf mm2tex} filter converts documents written in mm into \LaTeX\ using
the tm.sty style developed for writing AT\&T memoranda.  The filter
currently uses a combination of sed and awk scripts. 
\section{Status}
The filter
cannot handle all mm and troff macros but handles most of the commonly
used ones, passing the rest on through into the \LaTeX\ document where
they can be edited manually. Tables~\ref{tab:mmmacros} -
\ref{tab:strings} show which troff and mm features are currently translated.
\begin{table}[htp]
\input{mm2texmac1}
\caption{\label{tab:mmmacros}MM Macros Supported}
\end{table}
\begin{table}[htp]
\input{mm2texmac2}
\caption{\label{tab:mmmacros2}MM Macros Supported, con't}
\end{table}
\begin{table}[htp]
\input{mm2textr1}
\caption{\label{tab:troffmacros}Troff Macros Supported}
\end{table}
\begin{table}[htp]
\input{mm2textr2}
\caption{\label{tab:troffmacros2}Troff Macros Supported, con't}
\end{table}
\begin{table}[htp]
\input{mm2texsym}
\caption{\label{tab:specialsymb}Special Symbols}
\end{table}
\begin{table}[htp]
\input{mm2texnr}
\caption{\label{tab:numreg}MM Number Registers}
\end{table}
\begin{table}[htp]
\input{mm2texsr}
\caption{\label{tab:strings}MM String Registers}
\end{table}
\section{Limitations}
  \begin{enumerate}
    \item The sequences \verb|\*(Fg| and \verb|\*(Fg+1| are recognized
          but other added number is treated as 1.  The same is true
          for \verb|\*(Tb|.  \verb|\*(Rf| is recognized and
          autoincremented as normal.  \verb|.RS| and \verb|.RF|
          reference the previous value of Rf as expected.
    \item tabs not handled.
    \item The MM standard string registers may be interpolated (used
          in text) but not changed. 
    \item The MM number registers may be interpolated and those
          modified implicitly by mm macros (such as F is modified by
          .FS) are maintained but they may not be set explicitly.
  \end{enumerate}
\section{Usage}
The use the translator type
\begin{itemize}
\item[]
\verb|mm2tex |{\it mmfile}
\end{itemize}
If {\it mmfile\/} does not exist {\bf mm2tex} will try {\it mmfile}.mm.
The output is put into the file {\it mmfile}.tex.  
\section{Implementation}
The following
scripts are used by {\bf mm2tex} and must be accessible.
\begin{table}[htp]
\noindent
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{3}{|c|}{{\vrule height12pt width0pt depth3.5pt\Large
mm2tex Files}}\\ 
\hline\hline
\multicolumn{1}{|c|}{script}&\multicolumn{1}{c|}{type}&\multicolumn{1}{c|}{description}\\
\hline
mm2tex&      shell script&      main script\\
mm2tex.sh&   shell script&      raw filter\\
mm2tex.sed&  sed script&  pre filter\\
mm2tex.sed2&  sed script& post filter\\
mm2tex.awk&   awk script&     main translation\\
mm2tex.tex&   \TeX\ source&   this document\\
\hline
\end{tabular}
\end{center}
\medskip
\caption{\label{tab:scripts}mm2tex Files}
\end{table}
The mm2tex.sh script must be edited to know the path to the sed and
awk scripts.  mm2tex and mm2tex.sh must be in PATH.
\section{Availability}
The files are available from the author on request and are included
with the other tm.sty related files when they are distributed.
\makefacesignature{/home3/face}
%\noindent Atts.\\
%References\\
%Appendixes A-C\\
%Table 1

%\copyto{{\nobreak\topsep=0pt\partopsep=0pt\begin{tabbing}\nobreak%
%\hskip2in\=WW \=\kill%
%All Members Department 59114\\
%All Members Department 59112\\
%All Supervision Center 5911\\
%\copytoline{dgb}\\%Belanger
%\copytoline{jph}\\%Holtman
%\copytoline{jwh}\\%Hunt
%\copytoline{fml}\\%Lax
%\copytoline{jfm}\\%Maranzano
%copytoline{drr}\\%Ryan
%J.\ W.\ Foo		\>LC 0N-X00
%\end{tabbing}}}
%\bibliography{sim,rapid,tutorial,softengr}
%\appendices
%\input{glossary}
%\tableofcontents
%\listoffigures
%\coversheet
\end{document}
