% example of a letter using attletter.sty
%
\documentstyle{attletter}
\date{January 16, 1989}
\labelheight=1.5in
\makelabels
\phone{201 580-4428}\room{4N-E01}
\eaddress{tla@mozart.att.com}
\signature{Terry L Anderson\\Author of attletter.sty}
\begin{document}
\begin{letter}{Mr Really Somebody\\123 Fourth St\\Someplace, XX\quad 00000}
\opening{Dear Really}
This is a sample letter using the attletter style.  Note that I have
used the default \verb|\address| which is set to the address of my
location in the last few lines of attletter.sty.  I have added my
personal phone number which replaces that of the information operator
at my location and have added my room number and electronic address.

The height of each mailing label defaults to 1 inch in attletter
(compared to 2 inches in the standard letter.sty), but for this letter
I have changed it to 1.5 inches using \verb|\labelheight|.  I have
called \verb|makelabels| to illustrate this.  Labels can be printed on
or photocopied onto ``peel-off'' label sheets.

Many of the extensions to \LaTeX\ offered by tm.sty are also available
in attletter.sty.  These include 
\begin{alphenum}
\item \verb|\markbottom|
\item the additional types of enumerated lists
\item draft markings
\item the trademark macros
\item \UNIX\
\item \verb|\markedfootenote|
\end{alphenum}

Bitmaps can be used although there is not a special version of the
closing incorporating faces.  

This is a nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is a another nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is a another nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is a another nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is a another nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is a another nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

This is the last nonsense paragraph in order to make this letter long enough
to extend to a second sheet.

The default page style has been changed to {\it headings} so that
letters that extend onto additional pages have a heading at the top of
each subsequent page.  If you do not want this you can declare
\verb|\pagestyle{plain}| to have the style that was default in
standard letter.sty.

\closing{Sincerely,}
\end{letter}
\begin{letter}{Mrs. Anybody Else\\4321 Easy St\\Timbuktu, Mali}
\opening{Dear Mrs Else}
This is a second letter to illustrate that the attletter style like
the standard letter style supports multiple letters in a single file.
Each letter will have the letter head.  Labels if produced with
\verb|\makelabels| will be in two columns on as many sheets as
necessary, the number depending on the labelheight.
\closing{Yours Truly,}
\end{letter}
\end{document}

