%%% ====================================================================
%%% @LaTeX-file{
%%%   filename  = "diff12.tex",
%%%   version   = "1.2",
%%%   date      = "27-Jan-1995",
%%%   time      = "16:33:11 EST",
%%%   checksum  = "03031 442 3003 21900",
%%%   author    = "American Mathematical Society",
%%%   copyright = "Copyright (C) 1995 American Mathematical Society,
%%%                all rights reserved.  Copying of this file is
%%%                authorized only if either:
%%%                (1) you make absolutely no changes to your copy,
%%%                including name; OR
%%%                (2) if you do make changes, you first rename it
%%%                to some other name.",
%%%   address   = "American Mathematical Society,
%%%                Technical Support,
%%%                Electronic Products and Services,
%%%                P. O. Box 6248,
%%%                Providence, RI 02940,
%%%                USA",
%%%   telephone = "401-455-4080 or (in the USA and Canada)
%%%                800-321-4AMS (321-4267)",
%%%   FAX       = "401-331-3842",
%%%   email     = "tech-support@math.ams.org (Internet)",
%%%   supported = "yes",
%%%   keywords  = "amslatex, ams-latex",
%%%   abstract  = "This is part of the AMS-\LaTeX{} distribution. It
%%%                describes the differences between versions 1.1 and 1.2.",
%%%   docstring = "The checksum field above contains a CRC-16 checksum
%%%                as the first value, followed by the equivalent of
%%%                the standard UNIX wc (word count) utility output of
%%%                lines, words, and characters.  This is produced by
%%%                Robert Solovay's checksum utility.",
%%% }
%%% ====================================================================
\documentclass{amsdtx}
\usepackage{amsmath}

\title{Differences between \amslatex/ version 1.1 and \amslatex/
version 1.2}
\author{American Mathematical Society (Michael Downes)}
\date{January 1995}

\MakeShortVerb{\|}

%    Fake article-style maketitle since ams-doc-article class doesn't
%    exist yet
\makeatletter
\def\maketitle{\chapter*{\large\@title}%
  \section*{\large\normalfont\@author\\\@date}%
  \markboth{DIFFERENCES BETWEEN VERSIONS 1.1 AND 1.2}%
    {DIFFERENCES BETWEEN VERSIONS 1.1 AND 1.2}%
  \mbox{}% get some extra blank space
  \thispagestyle{plain}}
\makeatother

\renewcommand{\labelenumi}{(\theenumi)}

\begin{document}
\maketitle

\section{Overview}
\subsection{Transition to \LaTeXe{}}

Version 1.1 of \amslatex/ was designed to work with \latex/2.09, which
became obsolete as of June 1994; version 1.2 of \amslatex/ is designed
to work with the new \latex/, aka \LaTeXe{}. Generally speaking,
pre-existing documents that were written for \latex/2.09 can be
processed by current \latex/ through a `compatibility mode' that is
automatically entered when a document uses \cn{documentstyle} instead of
\cn{documentclass} (see below).

For generic \latex/ the emulation provided by \latex/'s compatibility
mode is extremely good---at least for well-formed documents that avoided
commands not documented in the \latex/ manual; the most common point of
failure is due to use of internal font commands such as \cn{fivrm} or
\cn{tenrm}, or to the loading of an option file that uses old internal
commands (font-related or otherwise). For third-party extensions such as
\amslatex/ the quality of the emulation in compatibility mode varies; at
the time of writing (18-Jan-1995) the emulation is fairly good for AMS
packages such as \pkg{amstex} or \pkg{amssymb}, but less good for the
AMS documentstyles \cls{amsart} and \cls{amsbook}; work to improve the
accuracy of the emulation is on-going.

In general, version 1.2 of \amslatex/ adheres as faithfully as possible
to the \latex/ conventions and command set described in the second
edition of \emph{\latex/: A document preparation system} (Lamport,
1994). For end users the changes lie primarily in the use of
\cn{documentclass} and \cn{usepackage} instead of \cn{documentstyle}
plus options, and in the changed syntax of font commands---e.g.,
|\textbf{...}| and |\mathbf{...}| instead of |{\bf...}|.

For example, to use the \pkg{amsbsy} package with a general document
class such as \cls{article}, you should no longer write
\begin{verbatim}
\documentstyle[amsbsy]{article}% obsolete
\end{verbatim}
but write instead
\begin{verbatim}
\documentclass{article}
\usepackage{amsbsy}
\end{verbatim}

For pre-existing documents running in compatibility mode the old forms
\cn{bf}, \cn{it}, etc., for the standard font commands now behave as
described in the first edition of the \latex/ manual. Depending on the
installation options used for your copy of \amslatex/ version 1.1, this
may be the `right' behavior for your existing \amslatex/ 1.1 documents;
if not, you will need to update the options list of the
\cn{documentstyle} command by adding \opt{newlfont}, for example:
\begin{verbatim}
\documentstyle[newlfont]{amsart}
\end{verbatim}

Internally the changes in version 1.2 of \amslatex/ were somewhat more
extensive; various parts were rewritten to take advantage of the most
recent version of the font selection scheme, for example, as well as the
new methods for managing packages and document classes, for defining
robust commands, for writing technical commentary in typesettable form,
and for providing installation scripts. In other words this version of
\amslatex/ is thoroughly integrated with current \latex/, not just
jury-rigged.

The user's guide for \amslatex/ 1.2 describes only the current behavior;
it doesn't attempt to describe the ways in which current behavior
differs from the behavior of previous versions (that's the purpose of
this document).

\subsection{Packages \pkg{amsfonts}, \pkg{amssymb} moved to the AMSFonts
distribution}

In version 1.2, \latex/ support for the use of fonts in the AMSFonts
distribution was moved out of the \amslatex/ distribution and into the
AMSFonts distribution, where it more properly belong. This means the
packages \pkg{amsfonts}, \pkg{amssymb}, \pkg{euscript}, and
\pkg{eufrak}, as well as associated font definition files. Please note
that certain Computer Modern fonts that were formerly found only%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\footnote{Here `only' is an oversimplification, in light of the
availability of (for example) the Sauter versions of the CM Metafont
source files.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
in the AMSFonts distribution are considered standard in \LaTeXe{}:
\fn{cmmib} and \fn{cmbsy} in sizes 5--9, and \fn{cmex} in sizes 7--9.
I.e., when you get a copy of \LaTeXe{} (and you cannot use \amslatex/
otherwise), those fonts should be included. If they are not, please
inquire to the source where you obtained \LaTeXe{}. If you obtained
\LaTeXe{} from CTAN, you can get Metafont source files for those fonts
in the directory \fn{/tex-archive/fonts/latex/mf}.

As \fn{.fd} files for those particular fonts are provided in the basic
\latex/ distribution, the AMSFonts distribution does not redundantly
provide them.

\subsection{Package \pkg{amstex} renamed to \pkg{amsmath}}

The \pkg{amstex} package has been renamed \pkg{amsmath}. The old name
`amstex' came from the fact that version 1.0 was a more or less straight
port into \latex/ of selected portions from the macro package \AmS-\tex/
written by Michael Spivak, but the historical origins of the package are
likely of less interest to the general usership---especially looking to
the future---than a straightforward declaration in the name `amsmath' of
the sort of features that are provided.

Another important reason for the name change was to make possible some
progress that would be out of the question (because of the impact on
existing documents) if the name `amstex' were retained. For example,
unlike \pkg{amstex}, \pkg{amsmath} doesn't automatically load the
\pkg{amsfonts} package, as that package was moved out to the AMSFonts
distribution as described above. If this change were not accompanied by
a name change, many existing documents would have to be updated by hand
(adding |\usepackage{amsfonts}|) in order to run without error. A frozen
version of \fn{amstex.sty} is provided for processing pre-existing
documents; it won't have active maintainence henceforth except for bug
fixes. For new documents authors should use the \pkg{amsmath} package
(not just because we advise it, but because it's a better package than
its predecessor in a number of ways).

\subsection{New documentclass \cls{amsproc}}

A new documentclass \cls{amsproc} for collections of articles, such as
the proceedings of a conference, is now provided in addition to
\cls{amsart} and \cls{amsbook}.

\section{Differences in the \pkg{amsmath} package}

Here we describe differences that users will see in the \pkg{amsmath}
package, as the successor of \pkg{amstex} version 1.1. In general,
commands and options that were dropped during the creation of the
\pkg{amsmath} package are retained in the frozen version of \pkg{amstex}
that is provided for compatibility mode use.

\begin{enumerate}

\item The options \opt{intlim}, \opt{nosumlim}, \opt{nonamelm},
\opt{righttag}, \opt{ctagsplt} for the \pkg{amstex} package are
superseded by more intelligible names \opt{intlimits},
\opt{nosumlimits}, \opt{nonamelimits}, \opt{reqno}, \opt{centertags} for
the \pkg{amsmath} package.

\item The \qc{\@} character is no longer used as a special command
prefix, except for the \pkg{amscd} package, and then only within the
\env{CD} environment. Practically speaking, this means primarily that it
is no longer necessary to use doubled |@@| to get a single printed @
character (e.g., in e-mail addresses).

\item The extensible arrow commands |@>>>| and |@<<<| are no longer
available outside of the \env{CD} environment of the \pkg{amscd}
package; alternative commands \cn{xleftarrow} and \cn{xrightarrow} are
provided by the \pkg{amsmath} package. The sub and superscripts are
given through an optional resp. mandatory argument, for example
\begin{ctab}{ll}
|X\xleftarrow[a]{b}Y|& $X\xleftarrow[a]{b}Y$\\
|X'\xrightarrow[\alpha]{\beta}Y'|& $X'\xrightarrow[\alpha]{\beta}Y'$
\end{ctab}

\item The command \cn{bold} has been superseded by the command
\cn{mathbf}, as the latter is now the standard name provided by \latex/
for this purpose.

\item The commands \cn{newsymbol}, \cn{frak}, \cn{Bbb} are no longer
available because the \pkg{amsfonts} package is not loaded by
\pkg{amsmath}. (In any case those commands now have new names---see the
\pkg{amsfonts} documentation for details.)

\item The \pkg{amsmath} package provides \cn{lvert} and \cn{rvert} for
$\vert$ symbols when they act as delimiters (compare \cn{langle},
\cn{rangle} for $\langle\ldots\rangle$ delimiters). This is to address
the problem that the \qc{\|} character is overloaded to represent
several different notations with \emph{different typographical
treatment}. Although this overloading is handled easily enough in
reading by the discriminatory powers of the knowledgeable reader, it
presents more difficult obstacles to the processing of electronic
documents by various computer tools. Commands \cn{lVert} and \cn{rVert}
are likewise provided in stead of \cn{\|} for paired-delimiter use of
the $\|$ symbol. Further discussion is found in the \amslatex/ user's
guide (\fn{amsldoc.tex}).

\item The \qc{\~} command is left unchanged by \pkg{amsmath}; the change
that makes it remove a redundant preceding or following space character
is now done only by AMS documentclasses (in version 1.1 this was done by
the \pkg{amstex} package).

\item The following rarely used commands are gone: \cn{accentedsymbol},
\cn{sphat}, \cn{spcheck}, \cn{sptilde}, \cn{spdot}, \cn{spddot},
\cn{spdddot}, \cn{spbreve}, \cn{@)))}, \cn{@(((}. They can be obtained
through the \pkg{amsxtra} package if necessary.

\item The six generalized fraction commands \cn{over},
\cn{overwithdelims}, \cn{atop}, \cn{atopwithdelims}, \cn{above},
\cn{abovewithdelims} are expressly forbidden by the \pkg{amsmath}
package; use of the recommended forms \cn{frac}, \cn{binom}, and
variants is now required, instead of merely recommended, and a new
command \cn{genfrac} has been added to fill in the access gaps that used
to exist. Use of the \cn{genfrac} command is discussed in the \amslatex/
user's guide (\fn{amsldoc.tex}).

Not only is the unusual syntax of the primitive \tex/ fraction commands
rather out of place in \latex/, but furthermore that syntax seems to be
solely responsible for one of the most significant flaws in \tex/'s
mathematical typesetting capabilities: the fact that the current
mathstyle at any given point in a math formula cannot be determined
until the end of the formula, because of the possibility that a
following generalized fraction command will change the mathstyle of the
\emph{preceding} material. As the side effects are a bit technical in
nature, they are discussed in \fn{technote.tex} rather than here.

\item The \cn{fracwithdelims} command is gone, as \cn{genfrac} provides
equivalent functionality (and a little more).

\item The optional argument of \cn{frac} that allowed changing the
thickness of the fraction line is gone. That functionality is now
provided only through \cn{genfrac}: the need to change the line
thickness is so rare in practice that it seems better not to burden the
ubiquitous \cn{frac} command with the somewhat time-consuming look-ahead
process required for an optional argument. I.e., this change will
contribute in a small way to making your documents run faster.

\item The commands \cn{lcfrac} and \cn{rcfrac} for left or right
alignment of continued fraction numerators are gone; they are replaced
by an optional argument of the \cn{cfrac} command.

\item A new command \cn{DeclareMathOperator} is provided for defining
new operator names:
\begin{verbatim}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\xlim}{x-lim}
\end{verbatim}
The |*| form makes the operator name produce limit-style sub and
superscripts, like \cn{max} or \cn{lim}. This is now the preferred
method, instead of using the old command \cn{operatorname} (which is,
however, retained for now). The command \cn{operatornamewithlimits} has
been renamed \cn{operatorname*}. These commands now can also be obtained
through a separate package \pkg{amsopn}, so it's not necessary to load
the entire \pkg{amsmath} package just to get that feature.

\item The environments \env{Sb} and \env{Sp} for multiline subscripts
have been replaced by a single command \cn{substack}, used as follows:
\begin{verbatim}
\sum_{\substack{first line\\second line}}
\end{verbatim}
This command can be used equally well in a subscript or superscript, and
provides better vertical positioning (as compared to \env{Sb}, \env{Sp})
if a multiline subscript is placed on the side, as perhaps for an
integral. There is also a slightly generalized alternative, an
environment \env{subarray} that allows you to specify either centering
or left alignment for the contents:
\begin{verbatim}
\sum_{\begin{subarray}{l}first line\\second line\end{subarray}}
\end{verbatim}

\item A command \cn{nobreakdash} is provided to suppress the possibility
of a linebreak after the following hyphen or dash. For example, if you
write `pages 1--9' as |pages 1\nobreakdash--9| then a linebreak will
never occur between the dash and the 9. You can also use
\cn{nobreakdash} to prevent undesirable hyphenations in combinations
like |$p$-adic|. For frequent use, it's advisable to make abbreviations,
e.g.,
\begin{verbatim}
\newcommand{\p}{$p$\nobreakdash}% for "\p-adic"
\newcommand{\Ndash}{\nobreakdash--}% for "pages 1\Ndash 9"
\end{verbatim}

\item The \env{[xx]align[at]} family of environments has been thoroughly
revised: now the \env{xalignat} environment is gone, as its function has
been merged into the \env{align} environment (it's no longer necessary
to specify how many side-by-side structures, as was required for
\env{xalignat}, because \env{align} automatically handles any number of
them). Similarly, the \env{xxalignat} environment has been replaced by
an environment \env{flalign} that doesn't require you to specify how
many side-by-side structures will be needed. Certain numbering problems
(presence of an equation number when it should be absent, or vice versa)
have also been cleared up.

\item The placement of equation numbers has been substantially improved
for \env{align}, \env{gather}, and \env{split}. Numbers will now
\emph{never} overlap on top of the equation body (as far as we can
ascertain), and they are much less likely to be shifted up or down
unnecessarily when there is actually adequate space to leave the number
in the normal place. In a few remaining cases, an equation number will
be shifted unnecessarily because technical complications make accurate
measurement of the available room too difficult; for those cases a
\cn{raisetag} command is provided that allows you to manually adjust the
vertical position of the equation number. (Thanks to David M. Jones for
his substantial [volunteer] work behind this change and the preceding
one.)

\item The seldom-used command \cn{oldnos} is gone; if you need it, you
should consult the \latex/ documentation on using fonts (e.g.,
\fn{fntguide.tex}) to find out how to construct an equivalent command.

\item A new \env{subequations} environment causes all numbered equation
environments within its scope to be numbered (4.9a) (4.9b) (4.9c) etc.,
if the preceding numbered equation was 4.8. A \cn{label} command
immediately after |\begin{subequations}| will produce a \cn{ref} of the
parent number `4.9', not `4.9a'. The counters used by the
\env{subequations} environment are |parentequation| and |equation| and
standard uses of \cn{addtocounter}, \cn{setcounter}, etc., are possible
with those counter names.

\item The \pkg{amsmath} package (unlike \pkg{amstex}) respects the
setting provided by the documentclass for putting equation numbers on
the right or the left. [Technical note: a documentclass that has
\opt{leqno} as the default should explicitly pass that option to the
\pkg{amsmath} package with \cn{PassOptionsToPackage} in order for it to
be effective.]

\item The \opt{fleqn} option for left-aligned instead of centered
equations is now supported (thanks to David M. Jones).

\end{enumerate}

\section{Differences in the \cls{amsart} and \cls{amsbook} document
classes}

\begin{enumerate}

\item As mentioned earlier, a companion document class \cls{amsproc} is
now provided for books that consist of a collection of articles, such as
the proceedings of a conference.

\item As the AMS document classes automatically load the \pkg{amsmath}
package, all changes described above for the \pkg{amsmath} package
affect the AMS classes also. Note in particular that the \qc{\@}
character is no longer a special command prefix, so to produce a printed
\qc{\@} in an e-mail address you should now write simply \qc{\@} instead
of |@@|.

\item The nonbreaking dash commands |@-|, |@--|, |@---| are no longer
available because \qc{\@} as a special command prefix has disappeared.
An alternative \cn{nobreakdash} command is provided in the \pkg{amsmath}
package, as described above.

\item The AMS document classes now take an option \opt{nomath} that
suppresses the automatic loading of the \pkg{amsmath} package. This is
sometimes helpful in converting a pre-existing document to AMS style if
you only want to do a quick conversion without attempting to
deal with any of the math.

\item The old environments \env{pf}, \env{pf*} are superseded by a
single \env{proof} environment. It takes an optional argument to specify
an alternative heading text.

\item Formerly \cn{small} produced the same typesize as
\cn{footnotesize} (8pt). Now they produce different sizes, normally 9
and 8 respectively. Furthermore, documentclass options \verb'8pt',
\verb'9pt', \verb'11pt', \verb'12pt' are now offered in addition to the
default \verb'10pt'. Also the range of typesize-changing commands is
filled out to include, below \cn{normalsize}, the following: \cn{small},
\cn{Small}, \cn{SMALL}, \cn{tiny}, \cn{Tiny}. \cn{footnotesize} and
\cn{scriptsize} are retained as synonyms of \cn{Small} and \cn{SMALL}
respectively. As it stands \cn{tiny} no longer produces the same
typesize as it did before (now 6 instead of 5), except in compatibility
mode. This might affect some existing documents if they are updated to
use \cn{documentclass} instead of \cn{documentstyle}. Similarly,
\cn{large} now produces 11pt if the base size is 10pt, where formerly it
produced 12pt. Finally, new commands \cn{larger} and \cn{smaller} are
provided for changing the typesize relative to the current size. These
commands take an optional integer argument to specify how many steps to
go up or down: \verb'\larger[2]' means go up two sizes, and
\verb'\larger' without an optional argument is the same as
\verb'\larger[1]'. The list of sizes is a standard progression of type
sizes and associated linespacing values as defined by the documentclass.
If you need finer control over type size or linespacing, consult the
\latex/ documentation about the \cn{fontsize} command.

\item The \cn{qed} symbol is now an open square, not a filled black
square, and it is positioned at the right margin, instead of at a fixed
horizontal distance from the preceding text.

\item The command \cn{rom}, for making numbers and punctuation
roman/upright in italic text, has been renamed \cn{upn} ``upright
punctuation or number''. It will in most cases suffice to leave this
refinement undone until a document is sent to a publisher for final
typesetting, and furthermore, certain publications at the AMS now use
special in-house italic fonts that have upright numbers and punctuation
built in, making the use of \cn{upn} unnecessary even then. For general
preprint-type use, this refinement is one that most users probably won't
care to bother with.

\end{enumerate}

\end{document}
